---
title: "Week 7 Data 605 Assignment"
author: "Tyler Baker"
date: "3/12/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
```

# Week 7 Assignment

## Important Distributions, Expected Value, Variance

1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to k. Let Y denote the
minimum of the Xi’s. Find the distribution of Y .

Answer
1. Since X1 up to Xn are mutually independent we know that 
P(X1 intersection X2) = P(X1)P(X2) for all Xi.
2. Since each Xi is uniformly distributed, we know that the probability of each l, where 1<=l<=k is 1/k.
3. I would like to think of each Xi as roll of k-sided dice. 
4. Thus, I think the minimum would be rolling a 1 on a k-sided die n times.
I think that would make the probability of that happening would be 1/((k^2)(n)).

2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).
This machine has a manufacturer’s expected lifetime of 10 years. This means
that we expect one failure every ten years. (Include the probability statements
and R Code for each part.).

Note
p = P(machine will fail within 10 years) = .1

p-1 = P(machine will not fail within 10 years) = .9

a. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)

Answer
Geometric distribution
P(X<=x) = 1 - (1-p)^x

P(X>x) = 1 - P(X<=x)

so,
P(X>8) = 1 - P(X<=8)
P(X>8) = 1 - (1-(1-.1)^8)
P(X>8) = 1 - (1 -.9^8)

```{r}
print(1 - (1 - (.9)^8))

p <- .1
q <- .9

pgeom(8,p,lower.tail = FALSE)
```
I'm not sure why these numbers are not matching up.

Expected value of Geometric distribution is 1/p.

```{r}
print(1/p)
```
Variance of geometric distribution is (1-p)/p^2
```{r}
print((1-p)/((p)^2))
```
So SD = sqrt(var(X))
```{r}
print(sqrt((1-p)/((p)^2)))
```


b. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential. 


lambda = l = .1
P(X>x) = 1 - P(X<=x)
and,
P(X<=x) = 1 - e^
so,
P(X>x) = 1 - (1 - e^(-.1*8))
```{r}
print(1 - (1 - exp(-.1*8)))
pexp(8, .1, lower.tail = FALSE)
```
The expected value is 1/ lambda.
```{r}
print(1/.1)
```
The variance is 1/ (lambda)^2.
So the standard deviation is
sqrt(1/lambda^2)
```{r}
print(sqrt(1/(.1)^2))
```

c. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years)

b(x, n, p) = (n choose x) * p^x * (1-p)^(n-x)

p = .1
n = 8
x = 0

Therefore,
```{r}
sol <- choose(8,0) * ((.1)^0) * ((.9)^(8-0))
print(1- sol)
pbinom(0, 8, .1, lower.tail=FALSE)
```
The Expected Value for a binomial distribution is n * p.
```{r}
print(8 * .1)
```
The standard deviation is sqrt(n * p * (1-p))
```{r}
sqrt(8 * .1 * .9)
```
d. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson. 

lambda is equal to the expected value which is equal to the expected number of successes. 
In this case, lambda = l = .8

The CDF is the sum from i=0 to x with the function (e^(-l)l^i/(i!))

However, a sum and an integral are essentially the same.

So I can just use the function (e^(-l)l^x/x!)

```{r}
integrand <- function(x) {(exp(-.8)*((.8)^x))/factorial(x)}
integrate(integrand, lower= 8, upper = 10)
```


```{r}
ppois(0, .8, lower.tail = FALSE)
```
I definitly did something wrong with my math above but not sure where I messed up.

The expected value again is lambda.

The variance is also lambda.

The standard deviation is sqrt(lambda).
```{r}
print(sqrt(.8))
```
